Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-01 paper/disc20clustering.tex

\maciek{Here, we need a good and broad description of what a component is.}

In this paper, we use the technique of maintaining \emph{(connected) components} of nodes, similar to previous approaches~\cite{repartition-disc}. A~\emph{component} is a subset of frequently communicating nodes \maciek{This is not a definition, just some intuitions. More formal needed.}.
We use this concept in both algorithms in this paper, but also in the lower bound, to incur high cost for any online algorithm that splits them.
An~algorithm is \emph{component respecting}
if it always keeps nodes that  belong to the same component collocated.
That is,
if the algorithm needs to move a node,
it moves the whole component containing the node.
%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
%This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
%such that frequently communicating nodes (from different clusters)
%are relocated to the same cluster before they inflict too much communication cost.
%A \emph{reconfiguration} (i.e.,\emph{ re-partitioning})
%is performed by migrating nodes between clusters,
%at the total cost of individual node migrations.
%The objective is to jointly minimize communication and migration costs while serving the sequence.

\maciek{Let's follow the SIDMA notation here. For example, a "single" node (the one that belongs to singleton component) is called "trivial" there.}

@@

@@

\maciek{TODO: rename everywhere: "adversary's components", i.e., all components in lower bounds should be called "ground sets"}

\maciek{The text before the proof should have two parts: a description of ground sets and then general intro to the proof.}

In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
For $k=2$, a constant competitive algorithm exists.
At the end of this section, we also
discuss an upper bound.

@@

The adversary constructs components of nodes based on the partitioning choices of a given online algorithm.
Adversary issues requests between pairs of nodes that belong to the same component.
If an online algorithm splits a component at any moment,
the adversary issues requests to the split pair until the algorithm collocates them.
We utilize the fact that the online algorithm does not know the components of the adversary, and it may split some components while evicting nodes.
On the other hand, $\OPT$ initially collocates all nodes of each component and never splits them again.

>>>>

Across this paper, we often refer to groups of communicating nodes.
As this concept is used differently in the lower bound and the upper bounds, we name them differently to avoid confusion.

\noindent
\textbf{Components.}
In our algorithms, a~\emph{component} is a subset of frequently communicating nodes.
We maintain all nodes of each component collocated in one cluster~\cite{repartition-disc}.
If we need to move a node, we move the whole component that contains it.
We maintain a balanced partition of our components, a similar concept to partition of integers into sets with equal sum (\maciek{TODO: cite}).
In contrast, our partition is time-varying: communicating components are merged into one, and we adjust the partition.

\noindent
\textbf{Ground sets.}
In the lower bound, a~\emph{ground set} is a group of nodes that start communicating if an algorithm splits them (places at different clusters).
Their role is to ensure that the algorithm maintains a partition of ground sets.
We emphasize that ground sets are revealed only upon their nodes are split.
\maciek{Other names: division, segmentation, alignment, formation}

\maciek{Needed? "An~algorithm is \emph{component respecting}
if it always keeps nodes that  belong to the same component collocated.
That is,
if the algorithm needs to move a node,
it moves the whole component containing the node."}
%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
%This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
%such that frequently communicating nodes (from different clusters)
%are relocated to the same cluster before they inflict too much communication cost.
%A \emph{reconfiguration} (i.e.,\emph{ re-partitioning})
%is performed by migrating nodes between clusters,
%at the total cost of individual node migrations.
%The objective is to jointly minimize communication and migration costs while serving the sequence.

We say that a component or a ground set is \emph{trivial}, if it contains a single node.

@@

At the end of this section, we discuss an asymptotically optimal upper bound.

@@

In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
To show it, clusters of capacity $k\geq 3$ are sufficient.
Note that for $k=2$, a lower bound of $3$ is known \cite{repartition-disc}.

@@

The adversary constructs ground sets in reaction to choices of a deterministic online algorithm.
Once we construct a ground set, it lasts until the end of the input sequence.
If the algorithm splits any pair of nodes of a ground set, the adversary issue requests to any pair of split nodes.

We construct the ground sets in a way that they always form a balanced partition of nodes.
This way, the optimal algorithm may move to the ground-set-respecting configuration, and serve all requests for free.
(This implies that the algorithm must eventually move to a ground-set-respecting configuration if it is competitive.)

\maciek{Finished here.}

We utilize the fact that the online algorithm does not know the components of the adversary, and it may split some components while evicting nodes.
On the other hand, $\OPT$ initially collocates all nodes of each component and never splits them again.
